---
title: "Shiny App - Decision Tree for predicting iris Species"
author: "ElisaRMA"
date: "8/31/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

The first step to built the app was to create a Machine Learning model. The objective was to create a model to predict the iris species based on their features. For this, the `caret` package was used along with the decision tree algorithm (`rpart`)

The iris data with the features used to build the model are as follows:
```{r}

data(iris)
head(iris)

```

## Algorithm workflow

The workflow for the creation of the algorithm was:

* Load `caret` package
* Partition of the data into training and testing
* Train the algorithm using the `rpart` method(decision tree) and cross validation
* Apply  model into the test set to verify the error rate
* Creation of the decision tree plot
* Apply the model into the new data given by the user

## Algorithm code
The code for the algorithm is below:
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(caret)

partition <- createDataPartition(y=iris$Species,p=0.7,list=FALSE)
train <- iris[partition,]
test <- iris[-partition,]

dtree <-  train(Species ~ .,data=train,method="rpart",
                trControl = trainControl(method = "cv"))
    
testpred <-  predict(dtree, newdata = test)
    
error  <- round(mean(testpred != test$Species), 2)

```


## The finished model 

```{r cars, echo=TRUE, message=FALSE, warning=FALSE}
library(rattle)
fancyRpartPlot(dtree$finalModel, uniform=TRUE)
```

## Input and Output of the app

The code presented before is embedded into the Shiny app along with the programming for appropriate interactivity. 

When values are inputted into the side panel of the app the model is applied to such values and a output of the predicted species is given.

An example of some values and its output is given below. 

```{r, echo=TRUE}
sl =1.5 ; sw =10 ; pl=15 ; pw = 7
predict(dtree, newdata=data.frame(Sepal.Length=sl, Sepal.Width=sw, 
                                  Petal.Length=pl, Petal.Width=pw))

```

